---
title: "Data-Driven Analysis of Attendance Trends in Los Angeles Dodgers Home Games"
author: "Preston Kilwien"
date: "2025-12-16"
output:
  pdf_document: default
  html_document: default
---
# Summary
This analysis uses the 2022 Dodgers home game dataset to identify what drives higher or lower attendance numbers. From the data analysis, I will turn in my findings to management so we can implement a strategy for consistent attendance numbers.
```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
# Load Libraries
```{r}
library(tidyverse) # data wrangling + visualization
library(lubridate) # easier date handling
```
# Read CSV file and quick look at sample rows
```{r}
dodgers_raw <- read.csv(
"/Users/bellevue/DSC 630/dodgers-2022.csv",
stringsAsFactors = FALSE
)
head(dodgers_raw)
```
# Check for missing values by column
```{r}
colSums(is.na(dodgers_raw))
```
```{r}
dodgers <- dodgers_raw
```
# Create only variables we actually need and verify structure
```{r}
dodgers <- dodgers %>%
mutate(
# Weekend vs weekday flag
weekend = ifelse(day_of_week %in% c("Friday", "Saturday", "Sunday"),
"Weekend", "Weekday"),
# Single promotion flag: Yes if any promotion exists
any_promo = ifelse(
  cap == "YES" | shirt == "YES" | fireworks == "YES" | bobblehead == "YES",
  "Yes", "No"
)
)
glimpse(dodgers)
```
# Attendance summary
```{r}
summary(dodgers$attend)
quantile(
dodgers$attend,
probs = c(0.05, 0.25, 0.5, 0.75, 0.95),
na.rm = TRUE
)
```
# Weekend Versus Weekday
```{r}
weekend_summary <- dodgers %>%
group_by(weekend) %>%
summarise(
games = n(),
avg_attend = mean(attend),
median_attend = median(attend)
)
weekend_summary
```
```{r}
ggplot(dodgers, aes(x = weekend, y = attend)) +
geom_boxplot() +
labs(
title = "Attendance by Weekend vs Weekday",
x = "Game Type",
y = "Attendance"
)
```
# Day of the Week
```{r}
dow_summary <- dodgers %>%
group_by(day_of_week) %>%
summarise(
games = n(),
avg_attend = mean(attend)
) %>%
arrange(desc(avg_attend))
dow_summary
```
```{r}
ggplot(dow_summary,
aes(x = reorder(day_of_week, avg_attend),
y = avg_attend)) +
geom_col() +
coord_flip() +
labs(
title = "Average Attendance by Day of Week",
x = "Day of Week",
y = "Average Attendance"
)
```
# Impact of Promotions 
```{r}
promo_summary <- dodgers %>%
group_by(any_promo) %>%
summarise(
games = n(),
avg_attend = mean(attend)
)
promo_summary
```
```{r}
ggplot(dodgers,
aes(x = any_promo, y = attend)) +
geom_boxplot() +
labs(
title = "Attendance: Promotional vs Non-Promotional Games",
x = "Any Promotion",
y = "Attendance"
)
```
# Weekday Promotions
```{r}
promo_weekday <- dodgers %>%
group_by(weekend, any_promo) %>%
summarise(
games = n(),
avg_attend = mean(attend),
.groups = "drop"
)
promo_weekday
```
```{r}
ggplot(promo_weekday,
aes(x = weekend, y = avg_attend, fill = any_promo)) +
geom_col(position = "dodge") +
labs(
title = "Promotion Impact by Day Type",
x = "Day Type",
y = "Average Attendance",
fill = "Promotion"
)
```
# Weather impact
```{r}
cor(dodgers$temp, dodgers$attend)
```
```{r}
ggplot(dodgers, aes(x = temp, y = attend)) +
geom_point(alpha = 0.7) +
geom_smooth(method = "lm", se = FALSE) +
labs(
title = "Attendance vs Temperature",
x = "Temperature",
y = "Attendance"
)
```
# Final Recommendations to Management:

1. Focus promotional efforts on weekday games, where baseline attendance is consistently lower than on weekends.

2. Use high-impact promotions such as giveaways and fireworks to increase midweek demand.

3. Offer bundled weekday ticket packages that include food or parking to reduce the overall cost.

4. Avoid relying on pricing alone to drive attendance. Use promotions to influence attendance on lower-demand dates.

5. Allocate promotional resources strategically to games where the potential for attendance growth is greatest.


